BivData Setup

### 2020-02-04 doc/Bivalent_Analysis.Rmd
work and notes

setwd("~./MLH1repo/")
#load Clean BivData
load(file="data/CleanBivData_1.28.20.RData")

#load other data files, MLH1
load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")#added batch18

@@

#remove all the non-used strains!!
Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
Curated_BivData$strain <- droplevels(Curated_BivData$strain)
#MOLF

#non-ordered factors, 
Curated_BivData$subsp <- factor(Curated_BivData$subsp, ordered = FALSE)
Curated_BivData$sex <- factor(Curated_BivData$sex, ordered = FALSE)
Curated_BivData$strain <- factor(Curated_BivData$strain, ordered = FALSE)

#all numbers, numbers
#all relavent factors unordered

#I'd like to merge the additional strains used for the male tests

@@

### Mixed Model Tests

@@

#### Mixed Model 1

@@

```{r M1_1CO, echo=FALSE}

#run Mixed models

#all pooled SC lengths
SCLength_all_.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=mouse.avs_4MM) 
#summary(Reduced.strain.HQ)

drop1(SCLength_all_.MM.M1, test="Chisq") # p = 0.0002045 ***

exactRLRT(SCLength_all_.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
as.data.frame(round(summary(SCLength_all_.MM.M1)$tTable, 6))

#1CO

#I think I need to re-calq the mouse averages for just 1CO
#mean.SC_1CO is mean 1CO

SC.Length_1CO_MM <- lme(mean.SC_1CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(SC.Length_1CO_MM)$tTable, 6 ) )
```

#### 2CO

#### 3CO

#### long.biv

@@

APRIL, consider moving these strain chunks to the bivData set up

@@

```{r WSB.SC, echo=FALSE, eval=FALSE}
#make the dataset
WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)

#plot the differences
wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
wsb.ploty

#make a table of the means and n ect
wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)
#can I plot these stats

@@

#### male strain comparisons

```{r SC.length.male_tests, echo=FALSE}

## run the tests to check following expectations

@@

@@

pairwise.t.test(Ozone, Month, p.adj = "bonf")

@@

@@

$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$

```{r MM_IFB.ABS, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

#remove the SKIVE male line which is na
mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$mean_IFD.2CO_ABS) | mouse.avs_4MM$mean_IFD.2CO_ABS==""), ]

lme_MM_IFD_ABS <- lme(mean_IFD.2CO_ABS ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),na.action=na.omit  )

as.data.frame( round(  summary(lme_MM_IFD_ABS)$tTable, 6 ) )
#p
#same missing values in obc

lme_MM_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ), na.action=na.omit )

as.data.frame( round(  summary(lme_MM_IFD_PER)$tTable, 6 ) )

#remove PWD, SKIVE and MSM
another.DF <- mouse.avs_4MM[mouse.avs_4MM$strain != "PWD",]
another.DF <- another.DF[another.DF$strain != "MSM",]
another.DF <- another.DF[another.DF$strain != "SKIVE",]

lme_MM_another_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= another.DF, random=list(strain=pdDiag(~sex) ), na.action=na.omit )

as.data.frame( round(  summary(lme_MM_another_IFD_PER)$tTable, 6 ) )

@@

#### Random Effects

These second set of models are made using the lmer()

$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$

I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure. The tables below are from anova( for the lmer model )

```{r MM.IFD, echo=FALSE}
#ugg what changed, FFF kntir is broken!!

IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(IFD_ABS.MM)

#thing1  <- summary(IFD_ABS.MM)
IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq")

rand_test_IFD.ABS.MM <- exactRLRT(IFD_ABS.MM)
rand_test_IFD.ABS.MM

IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(IFD_PER.MM)

Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
#Chis.sp_IFD.PER.MM

rand_test_IFD.PER.MM <- exactRLRT(IFD_PER.MM)
rand_test_IFD.PER.MM
```

Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER

@@

Dive deeper into the sex specific pattern for each strain.

these are example strains that show the 'unusual' results. show a raw plot, and norm plot and a scatter plot. Could the female bivalents with IFD under 30% be a different class of interference? (What does that mean?). I think the General pattern is that this rec pattern is not observed in males. That is there are rarely instances where 2 foci are positioned closer than 30% of the total SC length.

@@

@@

#make a table of the means and n ect

G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,

@@

@@

>>>>

setwd("~./MLH1repo/")
#load Clean BivData
load(file="data/CleanBivData_3.2.20.RData")
#load(file="data/CleanBivData_1.28.20.RData")

#load other data files, MLH1
load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")#added batch18

@@

@@

@@

#### HetC_M1

@@

Below's the code for Mixed model results -- try to organize them into a table or somethinf

```{r M1_many.metrics, echo=FALSE}
#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values

#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

#all pooled SC lengths
lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
anova(lmer.SCLength_all_MM.M1) # also F values

exactRLRT(lmer.SCLength_all_MM.M1)
as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
#can't get p values from this summary

#1CO
SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
drop1(SCLength_1CO_MM.M1, test="Chisq")
exactRLRT(SCLength_1CO_MM.M1)
as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))

#2CO
SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
drop1(SCLength_2CO_MM.M1, test="Chisq")
exactRLRT(SCLength_2CO_MM.M1)
as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))

#Long.Biv
SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)

drop1(SCLength_long.biv_MM.M1, test="Chisq")
exactRLRT(SCLength_long.biv_MM.M1)
as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))

#### HetC_M2

All fixed and all interactions

```{r M2_many.metrics, echo=FALSE, eval=FALSE}

#feed these 2 DFs into M2
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

#### HetC_M3

```{r HetC_M3, echo=FALSE}

@@

@@

@@

#### Male Poly

These are the models for male polymorphism, (I am re-thinking the tests I think)

Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**

```{r SC.length.male_tests, echo=FALSE}
## run the tests to check following expectations

@@

#Male.poly.Mouse.Table_BivData_4MM
#Long.biv_mouse_table_male.poly

@@

#pairwise.t.test(Ozone, Month, p.adj = "bonf")

@@

what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")

@@

```{r HetC_M1_IFD_many.metrics, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values

#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

#M1_IFD_2CO
lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
anova(lmer.IFD.PER_MM.M1)

exactRLRT(lmer.IFD.PER_MM.M1)
as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))

#M1_IFD.RAW_2CO
lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
anova(lmer.IFD.RAW_MM.M1) # also F values

exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

@@

I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure. The tables below are from anova( for the lmer model ). Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER

@@

Dive deeper into the sex specific pattern for each strain.  Below are code chunks which show the unusal sex specific results for IFD measures. The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD. The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.

For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?

@@

G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]

@@

#make a table of the means and n ect
G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,

@@

#calQ percent < 30%

oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]

#the na's are being counted (ughhhhhh)
#use the file thing instead

G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]

top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>% filter(IFD1_PER <= .3) #this makes DF
bottom.male <-  G.Bivalents.DF %>% filter(sex== "male")
ratio.male <- 18/1031 #2%

top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>% filter(IFD1_PER <= .3)
bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")
ratio.female <- 36/766  #5%

LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619

#uggh how does this not work ...

#more males have lower than 30% IFD?

LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
#0.7062663

#would Chi-sq be a good test to compare the ratios ect?

#the sum of these doesn't add

G.Bivalents.DF

@@

# old junk

### 2020-02-04 doc/Main_Report_Results.rmd
work and notes

load(file = "C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#2.2.20

load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")

@@

ToDo: add the random effect tests!!

@@

$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

@@

- ( Nested strain in subsp?)

@@

```{r mixed.model.data.setup, echo=FALSE, warning=FALSE}
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

@@

#reformate Molosinus subsp
#mean_co.HQ.mol
DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))

# For some reason I changed subsp to nums (wtf)
DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))

DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

@@

@@

```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
#mean_co.HQ  (this might remove some of the MOLF males)

@@

Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.

> Model 2.

@@

```{r MixedModel.2, echo=FALSE}
#reduced models
Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
#summary(Reduced.strain.HQ)

as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
#add a aption to this code chunk
#ranef(Reduced.strain.HQ)
```

@@

@@

```{r MixedModel.3, echo=FALSE}
#strain = fixed
lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#library(xtable)

>>>>

#load(file = "C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

#2.2.20
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")

@@

ToDo: change the Mixed models to lmer() so that I can confidantly test random effect (variance)

@@

I won't run the model above -- even tho it is logical and makes sense -- I can't make good estimates for random effects.

$$mouse .var CO ~=~ subsp * sex + rand(strain) + \varepsilon $$

@@

(Move this into the Methods)

@@

```{r mixed.model.data.setup, echo=FALSE, warning=FALSE}
# MOVE this into BIVDATA SETUP!!!

#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

@@

#add_subsp function re-writen

#mean_co.HQ.mol
#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
#        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))

# For some reason I changed subsp to nums (wtf)
#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
#        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))

@@

USE the bottom one

@@

```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2, eval=FALSE}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
#mean_co.HQ  (this might remove some of the MOLF males)

@@

Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.  
-->

> Model 2. THIS SHOULD BE THE MAIN MIXED MODEL --lmer

@@

```{r MixedModel.1, echo=FALSE, digits = 3}
#THIS IS THE NEW MIXED MODEL

#reduced models
MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
#summary(Reduced.strain.HQ)

drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***

exactRLRT(MLH1.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0

as.data.frame(round(summary(MLH1.MM.M1)$tTable, 6))
#add a aption to this code chunk
#ranef(Reduced.strain.HQ)
```

Above is the mixed model M1,

@@

All fixed effects with all interactions.

@@

```{r MixedModel.3, echo=FALSE}

#strain = fixed
lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#library(xtable)

### 2020-02-04 src/BivData_Setup.Rmd
work and notes

  #print(c(i, " the i") )

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

@@

>>>>

  #print(c(i, " the i") )
  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

@@

#remove na's
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]
